\documentclass[10pt,%
               %draft,%
               a4paper]{physics_notes}







\begin{document}

%%Front Page
\title{Classical Mechanics 2}
\author{St. Aidan's Physics Society}
\date{\today}
\maketitle

%%Contents
\tableofcontents
\newpage

\section{Preface}
This course is one of the harder aspects of L2 Physics. The reader will more than likely be lost or struggle in some derivations or explanations (Hopefully not due to our bad writing). You might even have to compromise on what you understand. However, you should at least appreciate where an idea comes from and its motivation. This helps your understanding but more importantly should interest you! Physics is \textit{hard}. You might like it because of the reward at the end of a problem or because its \textit{fundamentally incredible - Brain Cox}. If you like coffee or grapefruit, you could see that the problem itself is bittersweet and can be enjoyed in its own right.  Theoretical Physics 2 is where your preference might be tested. -- St. Aidan's PhysSoc\par



These notes are meant as a further addition to the lectures and notes provided by the lecturer. You will get the most out of these notes if you have a general idea of what its talking about.
\newpage

\section{Lagrangian Mechanics}
\input{"Sections/Lagrangian_Mechanics.tex"}


\section{Central Forces}
\input{"Sections/Central Forces.tex"}

\section{Hamiltonian Dynamics}
\input{"Sections/Hamiltonian_Dynamics.tex"}

\section{Rotating Coordinate Systems}
\input{"Sections/Rotations.tex"}

 




\end{document}